Covariant Phase Space via MultiSymplectic Geometry

### 2017-05-28 cpsviams.tex
First Version

\title{Covariant Phase Space via MultiSymplectic Geometry}
\author{
        Antonio Michele Miti \\
                Department of Mathematics and Physics\\
        Università Cattolica del Sacro Cuore\\
        via Musei X, Brescia xxx, \underline{Italy}
}
\date{\today}

@@

@@

\begin{abstract}
In this talk we detail the construction of the covariant phase space starting from multisymplectic framework of geometric classic field theories following essentially the paper written by forger romero. 
%Con lo stesso linguaggio si può descrivere più precisamente l'algoritmo di Peierls partendo dalle variazione di soluzioni sottomesse ad una variazione di lagrangiana.

@@

\section{Introduction}
Let's briefly recall in what consists the multisymplectic approach to mechanics of field systems in order to understand its relationship with the covariant phase space approach.

\begin{figure}
  \centering
  \includestandalone[mode=buildnew]{ms_landscape}
  \caption{multisymplectic landscape at glance}
\end{figure}

\begin{itemize}
 \item the starting point is the \emph{configuration bundle} 
 	\begin{displaymath}
 		E = \left( \pi_{M,E} : E\rightarrow M \right)
 	\end{displaymath}
 	the object encoding the kinematics.
 \item from that we have constructed the \emph{I jet bundle} applying the jet functor to $E$ 
 	\begin{displaymath}
 	J^1E = \left( \pi_{J^1 M,E} : J^1 E\rightarrow E \right)
 	\end{displaymath} 
 	that is an affine bundle over $E$.
 \item Then, we considered other two vector bundle derived from the first jet:
 	\begin{displaymath}
 	\AffDualJet  = \left( \pi_{ \AffDualJet ,E} : \AffDualJet  \rightarrow E \right)
 	\end{displaymath}
	twisted affine dual of $J^1E$ called \emph{"extended multiphase space"}.
	\begin{displaymath}
		\LinDualJet  = \left( \pi_{ \LinDualJet ,E} : \LinDualJet \rightarrow E \right)
	\end{displaymath}
	twisted linear dual of $J^1 E$ called \emph{"standard multiphase space"}. \footnote{It can be seen as the dual bundle of the vector bundle $\vec{J^1E}$ linear model of $J^1E$}
	The first one can be also seen as an affine line bundle over the second one.

\item It turns out that  $\LinDualJet $ carries a naturally defined multisymplectic form $\omega = -\ExtD \theta$ derived by a multicanonical m-form $\theta$. \\
	(This has been obtained pulling back the tautological m-form $\theta_\Lambda$ from $\bigwedge^m (E)$.) \\
	Thus, it can be considered the theoretical analogue of the cotangent bundle in ordinary geometric mechanics.

\item The dynamics can be implemented in two alternative way, fixing a further structure.\\
	In the Lagrangian picture is fixed a bundle map:
	\begin{displaymath}
		\Lagrangian : J^1 E \rightarrow \bigwedge^m (M)
	\end{displaymath}
	while in the Hamiltonian framework si fixed a section:
	\begin{displaymath}
		\Hamiltonian : \LinDualJet \rightarrow \AffDualJet
	\end{displaymath}

\item Both of them determine a \emph{Legendre map} computed as a fiber derivative.
	Mainly we are interested  in
	\begin{displaymath}
		\mathbb{F} \Lagrangian : J^1 E \rightarrow \AffDualJet		
	\end{displaymath}
	The two dynamics function can be considered equivalent under the condition of \emph{hyper-regularity}\cite[chapter, p.~215]{Gimmsy}.

\item Via $\mathbb{F} \Lagrangian$ we can pull-back the canonical forms on $J^1 E$ to give the \emph{Poincarè - Cartan forms}
	\begin{displaymath}
		\theta_\Lagrangian = \mathbb{F} \Lagrangian^\ast \theta \qquad \omega_\Lagrangian = - \diff \theta_\Lagrangian
	\end{displaymath}
	and via $\Hamiltonian$ we can pull-back the canonical forms on $\LinDualJet$ to give the \emph{De Donder - Weyl forms}
	\begin{displaymath}
		\theta_\Hamiltonian = \Hamiltonian^\ast \theta \qquad \omega_\Hamiltonian = - \diff \theta_\Lagrangian
	\end{displaymath}

\item At last, are introduced the \emph{Euler-Lagrange map}
	\begin{displaymath}
		D_\Lagrangian : J^2 E \rightarrow V^\TwistedLinear E
	\end{displaymath}
	 a bundle-morphism acting on holonomic section submanifolds $\varphi(M)$ as:
	\begin{displaymath}
		\left\langle D_\Lagrangian(\varphi, \partial \varphi , \partial^2 \varphi) \right\vert \left. V \right\rangle = (j^1 \varphi)^\ast \left( j^1V \lrcorner \omega_\Lagrangian \right)
	\end{displaymath}
	 a bundle-morphism when evaluated on a vertical field $V$ on $E$. \\
	And the \emph{DeDonder-Weyl map}
	\begin{displaymath}
		D_\Hamiltonian : J^1\left(\LinDualJet \right) \rightarrow V^\TwistedLinear \left(\LinDualJet \right)
	\end{displaymath}
	acting on holonomic section submanifolds $(\varphi, \pi)(M)$ as:
	\begin{displaymath}
		\left\langle D_\Hamiltonian(\varphi, \pi, \partial \varphi , \partial \pi) \right\vert \left. V \right\rangle = (\varphi, \pi)^\ast \left( V \lrcorner \omega_\Hamiltonian \right)
	\end{displaymath}	
	when evaluated on a vertical field $V$ on $\LinDualJet$.

\item	
	Imposing the vanishing of this map on an holonomic section gives a condition which corresponds to the well-know motion equations when rapresented in coordinate.

@@

\clearpage \newpage
\section{Covariant phase space approach}
The gist of the transition from the \emph{multi-symplectic approach} to the \emph{Covariant Functional Approach} - based on the \emph{Covariant Phase Space} - is basically to perform a change of perspective.\\
We have to move our attention from the bundles to the cross-sections on the bundles.

@@

	\begin{definition}[Space of kinematics (off-shell) configurations]\label{Def:ConfSpace}
		Is the, generally non-linear, infinte-dimensional manifold :
		\begin{displaymath}
			\Conf \coloneqq \Gamma^\infty(F,M)
		\end{displaymath}
	\end{definition}
%
	\begin{remark}
		Since the jet functor $j^1$ acts injectively on sections, we can equivalently consider any other holonomic jet sections
		\begin{displaymath}
			\Conf \simeq j^k \Gamma^\infty(F)
		\end{displaymath}
	\end{remark}	
%
	\begin{definition}[Space of Dynamics (on-shell) configurations]\label{Def:SolSpace}
		Is the subset of $\Conf$:
		\begin{displaymath}
			\Sol \coloneqq \ker(D_\cdot) \subset \Conf
		\end{displaymath}
		containing all the smooth solutions of the motion equations corresponding to the  dynamical operator $D_\Lagrangian$ or $D_\Hamiltonian$ depending on the framework considered.
	\end{definition}

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the section of a suitable vector bundle composing with the jet functor $j$.

\begin{equation}
	D_\Lagrangian[\; ] = D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[ \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right)
\end{equation}

\begin{equation}
	D_\Hamiltonian[\; ] = D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[ (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{equation}

The target base is a vector space, more precisely a frechet space \cite[Cap.2]{Ban} and makes sense to ask whetever $D_\cdot [\phi]$ is equal to the 0-section.

Again, we would like to employ a collective notation to denote the motion equation operators $D_\Lagrangian$ and $D_\Hamiltonian$:
\begin{notation}
	Since $D_\Lagrangian$ and $D_\Hamiltonian$ depend on different order jet bundle over $F$, can not be introduced directly a common notation on the multi-symplectic ground but we have to pass to the functional framework:
\begin{displaymath}
	D[\; ] = \; \Gamma^\infty(F) \ni \phi \quad \mapsto \quad D[ \phi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear F \right)\right)
\end{displaymath}	
in other words, $D[\phi](x) \in V^\TwistedLinear_{\phi(x)}F$ is a twisted linear one-form.
\\
	The evaluation of operator $D$, when applied to $\phi$, on a test vector field $V \in \Gamma(\phi^\ast(VF))$, results:
	\begin{displaymath}
		\left\langle D[\phi] \right\vert \left. V \right\rangle =
		\phi^\ast \left( i_V \omega \right) 
		\begin{cases}
			(j^1 \varphi)^\ast \left( i_{j^1V} \: \omega_\Lagrangian \right) \\
			(\varphi, \pi)^\ast \left( i_V \: \omega_\Hamiltonian \right)
		\end{cases}              
        \in \Omega^m(M)
	\end{displaymath}

	where has been employed the following correspondences:
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\phi \in \Conf$		&	$j^1 \varphi \; , \quad \varphi \in \Gamma^\infty(E)$		&	$(\varphi, \pi)$	\\
			$V \in \Gamma(\phi^\ast V F )$ & $j^1 V$ & $V$ \\
			$\omega$ & $\omega_\Lagrangian$ & $\omega_\Hamiltonian$ \\
			\hline
		\end{tabular}
	\end{center}

	Noticing the recurring scheme we will also denote $	W = V^\TwistedLinear F $
\end{notation}

The remarkable feature of the $\Sol$ space is that it carries a naturally defined pre-symplectic form $\Omega$. Natural beside the fact that its very structure, as a set, is expicitly determined by the fixed Lagrangian or Hamiltonian.\\
Therefore, is the candidate to  be the field-theoretic (Covariant) phase space.

Since $\Conf$ and $\Sol$ are in general not linear, we have to regard them as a manifold.
Hence, we have to specify conveniently the structure of their tangent space in order to make clair the preceeding claim.

The first thing to point out is the following chain of inequalities:
\begin{equation}
	\Gamma^\infty_c\left(\phi^*(VF)\right) \; \subseteq \; T_\phi\Sol \; \subset \; T_\phi\Conf \; \subseteq \; \Gamma^\infty\left( \phi^\ast (VF) \right)
\end{equation}

\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{Pictures/variations.png}
%\caption{\label{fig:variations}Covariant tangent fields as regular variations.}
\end{figure}

This condition stems from the fact that we want to interpret regular curves on $\Conf$ as smooth variations of field configurations. \\
Therefore, a parametrized regular curve as to be regarded as a family of diffeomerphisms, the flow generated by a vertical tangent field on $F$.\\
The pull-back are there to remark that $T_\phi \Conf$ consists of the infinitesimal smooth variations of $\phi$ thus are vertical field of $F$ \underline{along} the section $\phi$.
\begin{notation}
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\delta \phi \in T_\phi \Conf$		&	$\delta \varphi$		&	$(\delta \varphi, \delta \pi)$ \\
			\hline
		\end{tabular}
	\end{center}
\end{notation}

According to this interpretation, $T_\phi\Sol$ should consists of the generators of those variations that maps, at least infinitesimally, a solution $\phi\in\Sol$ in another solution.
Rigourosly, this can be expressed in the following definition:

	\begin{definition}[Tangent space to the Space of Solutions]\label{Def:TangentSol}
		\begin{displaymath}
			T_\phi \Sol \coloneqq \ker(J[\phi]_\cdot) \qquad \phi \in \Sol
		\end{displaymath}
	\end{definition}
In other words, it consists of the solutions of the Jacobi equations.

\clearpage \newpage
\section{Jacobi Operator}\label{Sec:JacobiOperator}
Intuitively, the Jacobi equation are the formal linearizazion of the motion equation around a solution $\phi$.\\
Precisely this can be encoded in the following operator:

	\begin{definition}[Jacobi operator pertaining to $D$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp} is the bundle map over $F$:
		\begin{displaymath}
			J(\phi) \; : \; \phi^\ast(VF) \rightarrow	\phi^\ast(V^\TwistedLinear F) 
			\quad s.t. \quad  \left(J(\phi)\right) V
			= \sigma \left( \left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda](x) \right\rvert_{\lambda=0} \right)
		\end{displaymath}
		where $V\in (\phi^\ast(VF))_{\phi(x)} = V_{\phi(x)}F$ and $\phi_\lambda$ is a regular variation generated by a $\delta\phi$ such that $\delta\phi_x = V$ and
		$\sigma$ is the canonical bundle-morphism
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
	\end{definition}
	As a bundle-map, $J$ can be naturally interpreted functionally as
	\begin{displaymath}
		\left(J(\phi)\right) [\: ] : 
		\Gamma^\infty\left(\phi^\ast(VF)\right) \rightarrow	
		\Gamma^\infty \left( 
		\phi^\ast(V^\TwistedLinear F)
		\right)
	\end{displaymath}
%
\begin{figure}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/Jacobi.png}
\caption{\footnotesize \label{fig:jacobi}$ \left(J(\phi)\right) [ \delta\phi ]$ and $D$ target on the same space.}
\end{figure}
%
Since $D[\phi_\lambda](x)$ is a smooth curve on $ V^\TwistedLinear F $ that is vertical w.r.t. $M$, the derivative involved has to be intended as the tangent velocity vector, thus $\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]$ is the vertical vector field associated to the infinitesimal variation.
Exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\diff}{\diff \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\diff}{\diff \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D_\cdot[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$, 
		thus the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.
%
The operator is well defined since
\begin{displaymath}
	\phi \in \Sol \quad \Rightarrow \quad D[\phi] \subset 0(E)
\end{displaymath}

>>>>

\title{Covariant Phase Space via MultiSymplectic Geometry}
\author{
	International Ph.D. Program in "Science" \\
	"Differential geometry and applications to modern physics" \\
	Department of Mathematics and Physics\\
	Università Cattolica del Sacro Cuore\\
	via Musei X, Brescia xxx, \underline{Italy}
}

@@

\usepackage[italian,english]{babel}
\usepackage[utf8]{inputenc}

@@

\begin{abstract}
In this short talk we present the results contained in the paper \emph{"Covariant Poisson brackets in geometric field theory"} written by M. Forger and S. V. Romero \cite{forgeromero}.
In particular we focus our attention on the construction of the \emph{Covariant Phase Space} starting from the multi-symplectic approach to geometric classic field theories.
%Con lo stesso linguaggio si può descrivere più precisamente l'algoritmo di Peierls partendo dalle variazione di soluzioni sottomesse ad una variazione di lagrangiana.

@@

\section{Multisymplectic landscape at glance}
Let's briefly recall in what consists the multisymplectic approach to mechanics of field systems in order to understand its relationship with the covariant phase space approach.

\begin{center}
  \includestandalone[mode=buildnew]{Pictures/Figure_ms_landscape}
\end{center}

\begin{itemize}
 \item the starting point is the \emph{configuration bundle} 
 	\begin{displaymath}
 		E = \left( \pi_{M,E} : E\rightarrow M \right)
 	\end{displaymath}
 	the object encoding the kinematics.
 \item from that we have constructed the \emph{I jet bundle} applying the jet functor to $E$ 
 	\begin{displaymath}
 	J^1E = \left( \pi_{J^1 M,E} : J^1 E\rightarrow E \right)
 	\end{displaymath} 
 	that is an affine bundle over $E$.
 \item then, we considered other two vector bundle derived from the first jet bundle. 
 	The twisted affine dual of $J^1E$ 
 	\begin{displaymath}
 	\AffDualJet  = \left( \pi_{ \AffDualJet ,E} : \AffDualJet  \rightarrow E \right)
 	\end{displaymath}
 	called \emph{"extended multiphase space"} and th twisted linear dual of $J^1 E$
	\begin{displaymath}
		\LinDualJet  = \left( \pi_{ \LinDualJet ,E} : \LinDualJet \rightarrow E \right)
	\end{displaymath}
	called \emph{"standard multiphase space"} \footnote{$\LinDualJet$ can be seen as the dual bundle of the vector bundle $\vec{J^1E}$ linear model of $J^1E$}.\\
	The first one can be also seen as an affine line bundle over the second one.

\item It turns out that  $\LinDualJet $ carries a naturally defined multi-symplectic form $\omega = -\ExtD \theta$ derived by a multicanonical m-form $\theta$, obtained by pulling back the tautological m-form $\theta_\Lambda$ from $\bigwedge^m (E)$.\\
	Thus, it can be considered the theoretical analogue of the cotangent bundle in ordinary geometric mechanics.

\item Dynamics can be implemented in two different way by fixing a further structure.\\
	In the Lagrangian picture has to be fixed a bundle map:
	\begin{displaymath}
		\Lagrangian : J^1 E \rightarrow \bigwedge^m (M)
	\end{displaymath}
	, called \emph{Lagrangian density}, 	while in the Hamiltonian framework has to be fixed a section:
	\begin{displaymath}
		\Hamiltonian : \LinDualJet \rightarrow \AffDualJet
	\end{displaymath}
	called \emph{Hamiltonian density}.

\item Both of them determine a \emph{Legendre map}, computed as a fiber derivative, which allows us to move from one framework to another.
	Mainly we are interested  in the following:
	\begin{displaymath}
		\mathbb{F} \Lagrangian : J^1 E \rightarrow \AffDualJet		
	\end{displaymath}
	The two dynamics function $\Lagrangian$ and $\Hamiltonian$ can be considered equivalent under the condition of \emph{hyper-regularity}\cite[chapter, p.~215]{Gimmsy}.

\item Via $\mathbb{F} \Lagrangian$ we can pull-back the canonical forms on $J^1 E$ to give the \emph{Poincaré - Cartan forms} giving:
	\begin{displaymath}
		\theta_\Lagrangian = \mathbb{F} \Lagrangian^\ast \theta 
		\qquad 
		\omega_\Lagrangian = - \diff \theta_\Lagrangian
	\end{displaymath}
	Via $\Hamiltonian$ we can pull-back the canonical forms on $\LinDualJet$ to give the \emph{De Donder - Weyl forms} giving:
	\begin{displaymath}
		\theta_\Hamiltonian = \Hamiltonian^\ast \theta 
		\qquad 
		\omega_\Hamiltonian = - \diff \theta_\Lagrangian
	\end{displaymath}

\item At last, can be introduced two map in order to encoded the field equations.
	In the Lagrangian framework we consider the \emph{Euler-Lagrange map}
	\begin{displaymath}
		D_\Lagrangian : J^2 E \rightarrow V^\TwistedLinear E
	\end{displaymath}
	 a bundle-morphism acting on holonomic section submanifolds $\varphi(M)$ as:
	\begin{displaymath}
		\left\langle D_\Lagrangian(\varphi, \partial \varphi , \partial^2 \varphi) \right\vert \left. V \right\rangle = (j^1 \varphi)^\ast \left( j^1V \lrcorner \omega_\Lagrangian \right)
	\end{displaymath}
	 a bundle-morphism when evaluated on a vertical field $V$ on $E$.\\
	In the Hamiltonian framework we consider the \emph{DeDonder-Weyl map}
	\begin{displaymath}
		D_\Hamiltonian : J^1\left(\LinDualJet \right) \rightarrow V^\TwistedLinear \left(\LinDualJet \right)
	\end{displaymath}
	acting on holonomic section submanifolds $(\varphi, \pi)(M)$ as:
	\begin{displaymath}
		\left\langle D_\Hamiltonian(\varphi, \pi, \partial \varphi , \partial \pi) \right\vert \left. V \right\rangle = (\varphi, \pi)^\ast \left( V \lrcorner \omega_\Hamiltonian \right)
	\end{displaymath}	
	when evaluated on a vertical field $V$ on $\LinDualJet$.

\item	
	Imposing the vanishing of this map on an holonomic section gives a condition which corresponds to the well-know motion equations when represented in local coordinate charts.

@@

\clearpage \newpage
\section{Covariant phase space approach}
The gist of the transition from the \emph{multi-symplectic approach} to the \emph{(Covariant) Functional Approach} - based on the \emph{Covariant Phase Space} - is basically to perform a change of perspective.\\
We have to move our attention from the bundles to the cross-sections on the bundles.

@@

	\begin{definition}[Space of kinematics (off-shell) configurations]\label{Def:ConfSpace}
		Is the, generally non-linear, infinte-dimensional manifold :
		\begin{displaymath}
			\Conf \coloneqq \Gamma^\infty(F,M)
		\end{displaymath}
	\end{definition}
%
	\begin{remark}
		The jet functor $j^1$ acts injectively on smooth cross sections
		\begin{displaymath}
			\Conf \simeq j^k \Gamma^\infty(F)
		\end{displaymath}
		so, we can equivalently consider holonomic jet sections of any order as the element of the configuration space.
	\end{remark}	
%
	\begin{definition}[Space of Dynamics (on-shell) configurations]\label{Def:SolSpace}
		Is the subset of $\Conf$:
		\begin{displaymath}
			\Sol \coloneqq \ker(D_\cdot) \subset \Conf
		\end{displaymath}
		containing all the smooth solutions of the motion equations corresponding to the  dynamical operator $D_\Lagrangian$ or $D_\Hamiltonian$ depending on the framework considered.
	\end{definition}

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the section of a suitable vector bundle composing with the jet functor $j$.
%
\begin{align*}
	D_\Lagrangian[\; ] &= 
	D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[ \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right) \\
	D_\Hamiltonian[\; ] &= 
	D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[ (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{align*}
%
The target base is a vector space, more precisely a frechet space \cite[Cap.2]{Ban} and makes sense to ask whetever $D_\cdot [\phi]$ is equal to the 0-section.

Again, we would like to employ a collective notation to denote the motion equation operators $D_\Lagrangian$ and $D_\Hamiltonian$:
\begin{notation}
	Since $D_\Lagrangian$ and $D_\Hamiltonian$ depend jet bundle over $F$ of different order, can not be introduced directly a common notation on the multi-symplectic ground but we have to pass to the functional framework:
\begin{displaymath}
	D[\; ] = \; \Gamma^\infty(F) \ni \phi \quad \mapsto \quad D[ \phi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear F \right)\right)
\end{displaymath}	
in other words, $D[\phi](x) \in V^\TwistedLinear_{\phi(x)}F$ is a twisted linear one-form.
\\
	The evaluation of operator $D$, when applied to $\phi$, on a test vector field $V \in \Gamma(\phi^\ast(VF))$, results:
	\begin{displaymath}
		\left\langle D[\phi] \right\vert \left. V \right\rangle =
		\phi^\ast \left( i_V \omega \right) 
		\begin{cases}
			(j^1 \varphi)^\ast \left( i_{j^1V} \: \omega_\Lagrangian \right) \\
			(\varphi, \pi)^\ast \left( i_V \: \omega_\Hamiltonian \right)
		\end{cases}              
        \in \Omega^m(M)
	\end{displaymath}

	where has been employed the following correspondences:
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\phi \in \Conf$		&	$j^1 \varphi \; , \quad \varphi \in \Gamma^\infty(E)$		&	$(\varphi, \pi)$	\\
			$V \in \Gamma(\phi^\ast V F )$ & $j^1 V$ & $V$ \\
			$\omega$ & $\omega_\Lagrangian$ & $\omega_\Hamiltonian$ \\
			\hline
		\end{tabular}
	\end{center}	
	%
	Noticing the recurring scheme we will also denote $	W = V^\TwistedLinear F $.
\end{notation}

Remarkable feature of the $\Sol$ space is that it carries a naturally defined pre-symplectic form $\Omega$. Natural beside the fact that its very structure, as a set, is expicitly determined by the fixed Lagrangian or Hamiltonian.\\
Therefore, is the candidate to  be the field-theoretic (Covariant) phase space.

\begin{center}
\includegraphics[width=0.49\textwidth]{Pictures/variations.png}
%\caption{\label{fig:variations}Covariant tangent fields as regular variations.}
\end{center}

Since $\Conf$ and $\Sol$ are in general not linear, we have to regard them as a manifold.
Hence, we have to specify conveniently the structure of their tangent space in order to make clear the previous claim.

The first thing to point out is the following chain of inequalities:
\begin{displaymath}
	\Gamma^\infty_c\left(\phi^*(VF)\right) \; \subseteq \; T_\phi\Sol \; \subset \; T_\phi\Conf \; \subseteq \; \Gamma^\infty\left( \phi^\ast (VF) \right)
\end{displaymath}
%
This condition stems from the fact that we want to interpret regular curves on $\Conf$ as smooth variations of field configurations. \\
Therefore, a parametrized regular curve as to be regarded as a family of diffeomerphisms, the flow generated by a vertical tangent field on $F$.\\
The pull-back are there to remark that $T_\phi \Conf$ consists of the infinitesimal smooth variations of $\phi$ thus are vertical field of $F$ \underline{along} the section $\phi$.
\begin{notation}
	\begin{center}
		\begin{tabular}{|c|c|c|}
			\hline
			 & Lagrangian picture & Hamiltonian picture \\
			\hline
			$\delta \phi \in T_\phi \Conf$		&	$\delta \varphi$		&	$(\delta \varphi, \delta \pi)$ \\
			\hline
		\end{tabular}
	\end{center}
\end{notation}

According to this interpretation, $T_\phi\Sol$ should consists of the generators of those variations that maps, at least infinitesimally, a solution $\phi\in\Sol$ in another solution.\\
In other words, it consists of the solutions of the \emph{Jacobi equations}.
Rigourosly, this can be expressed in the following definition:
	\begin{definition}[Tangent space to the Space of Solutions]\label{Def:TangentSol}
		on a solution $\phi \in \Sol$, is the vector space
		\begin{displaymath}
			T_\phi \Sol \coloneqq \ker(J[\phi]_\cdot)
		\end{displaymath}
		where $J$ is the so-called \emph{Jacobi Operator}.
	\end{definition}

\section{Jacobi Operator}\label{Sec:JacobiOperator}
Intuitively, the Jacobi equation are the formal linearization of the motion equation around a solution $\phi$.\\
Precisely they can be encoded in the following operator:
%
	\begin{definition}[Jacobi operator pertaining to $D$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp} is the bundle map over $F$:
		\begin{displaymath}
			J(\phi) \; : \; \phi^\ast(VF) \rightarrow	\phi^\ast(V^\TwistedLinear F) 
			\quad s.t. \quad  \left(J(\phi)\right) V
			= \sigma \left( \left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda](x) \right\rvert_{\lambda=0} \right)
		\end{displaymath}
		where $V\in (\phi^\ast(VF))_{\phi(x)} = V_{\phi(x)}F$ and $\phi_\lambda$ is a regular variation generated by a $\delta\phi$ such that $\delta\phi_x = V$ and
		$\sigma$ is the canonical bundle-morphism \footnote{see \ref{Section:sigma} for further details.}
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
	\end{definition}
	%
	\begin{remark}
		Being $J$ a bundle morphism, it can be naturally interpreted functionally as
		\begin{displaymath}
			\left(J(\phi)\right) [\: ] : 
			\Gamma^\infty\left(\phi^\ast(VF)\right) \rightarrow	
			\Gamma^\infty \left( 
			\phi^\ast(V^\TwistedLinear F)
			\right)
		\end{displaymath}
	\end{remark}
%
  \begin{minipage}[r]{0.5\textwidth}
    \includegraphics[width=0.85\textwidth]{Pictures/Jacobi.png}
  \end{minipage}\hfill
  \begin{minipage}[l]{0.5\textwidth}
	Note that $ \left(J(\phi)\right) [ \delta\phi ]$ and $D$ target on the same space.
  \end{minipage}

Since $D[\phi_\lambda](x)$ is a smooth curve on $ V^\TwistedLinear F $, that is also vertical w.r.t. $M$, 
the derivative involved has to be intended as the tangent velocity vector,
thus $\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]$ is the vertical vector field associated to the infinitesimal variation.\\
Exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$.
		Hence the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.

Furthermore, since
\begin{displaymath}
	\phi \in \Sol \quad \Rightarrow \quad D[\phi] \subset 0(E)
\end{displaymath}
the action of $\sigma$ is legit and operator $J$ results well-defined.
%
